---
title: "Getting Started with sanityzeR"
description: >
  Learn how to get started with the basics of sanityzeR.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with sanityzeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## The goal of sanityzeR

Data scientists often need to remove or redact Personal Identifiable
Information (PII) from their data. This package provides utilities to
spot and redact PII from r data frames/Tibbles.

PII can be used to uniquely identify a person. This includes names,
addresses, credit card numbers, phone numbers, email addresses, and social
security numbers, and therefore regulatory bodies such as the European Union’s
General Data Protection Regulation (GDPR) and the California Consumer 
Privacy Act (CCPA) require that PII be removed or redacted from data sets 
before they are shared an further processed.

## Functionalities

This document introduces you to the fundamental tools of sanityzeR and shows 
you how to apply them with data frames.

There are three functions in this package including data cleaning, credit 
card number handling, and email address handling.


## Installation

You can install the development version of sanityzeR from
[GitHub](https://github.com/) with:

```{r install, eval=FALSE, message=FALSE}
# install.packages("devtools")
devtools::install_github("UBC-MDS/sanityzeR")
```

## Setup

```{r setup}
library(sanityzeR)
```

### Create a dummy dataframe

```{r createDF}
df <- data.frame(
 Name = c("My email address is 123456abcd@yahoo.com and zzzzz123@yahoo.mail Thank you.",
                          "Bill for: 4556129404313766",
                          "Maria",
                          "Ben",
                          "Tina"),
                 Age = c(23, 41, 32, 16, 26)
)
df
```

### Create Spotters with redacted function

The following spotters will replace the detected PII with a fixed string.

```{r createSpotterRedact}
spotter_1_r <- list(redact_email,FALSE,"EMAILADDRS")
spotter_2_r <- list(redact_creditcardnumber,FALSE,"CREDITCARD")
spotters_redacted <- list(spotter_2_r,spotter_1_r)
```

### Create Spotters with hash function

The following spotters will replace the detected PII with a hash.

```{r createSpotterHash}
spotter_1_h <- list(redact_email,TRUE,0)
spotter_2_h <- list(redact_creditcardnumber,TRUE,0)
spotters_hashed <- list(spotter_2_h,spotter_1_h)
```

## Clean PII with redacted

Replacing the PII with a fixed string.

```{r cleanPIIwithRedacted}
clean_data_frame(df, spotters_redacted)
```

## Clean PII with hash

Replacing the PII with a hash.

```{r cleanPIIwithHash}
clean_data_frame(df, spotters_hashed)
```
